Parameter estimation

### 2023-10-05 SFE_article_Parameter_Estimation/Article_Model_Parameters.tex
05.10

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. Use 1, 2 and 3. TO be removed later
%\paperwidth=\dimexpr \paperwidth + 6cm\relax
%\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
%\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{graphicx,float}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
This study aimed to investigate the supercritical extraction process of the caraway oil from caraway seeds. A distributed-parameter model is used to describe the fluid-solid extraction process with carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and an initial state estimate. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. The missing parameters which describe the extraction kinetic are obtained by solving the maximum likelihood estimation on experimental yield data under the normal error assumption. A set of laboratory experiments was performed in a partially filled extractor with a fixed bed operated under multiple constant operating conditions: $40~^\circ C$ / $200$ bar, $50~^\circ C$ / $200$ bar, $40~^\circ C$ / $300$ bar, and $50~^\circ C$ / $300$ bar.

@@

\clearpage \appendix \label{appendix}
\section{Appendix} 
\subsection{Thermodynamic}
\subfile{Sections/Thermodynamic_details}

>>>>

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. Use 1, 2 and 3. TO be removed later
%\paperwidth=\dimexpr \paperwidth + 6cm\relax
%\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
%\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{graphicx}
\usepackage{float}
\restylefloat{table}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
This study aimed to investigate the supercritical extraction process of the caraway oil from caraway seeds. A distributed-parameter model is used to describe the fluid-solid extraction process with carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, and an initial state estimate. The missing model parameters are obtained by solving the maximum likelihood estimation based on experimental yield data under the normal error assumption. A set of laboratory experiments was performed in a partially filled extractor with a fixed bed operated under multiple constant operating conditions: $40~^\circ C$ / $200$ bar, $50~^\circ C$ / $200$ bar, $40~^\circ C$ / $300$ bar, and $50~^\circ C$ / $300$ bar.

@@

\clearpage \appendix \label{appendix}
\section{Appendix} 
\subsection{Thermodynamic}
\subfile{Sections/Qubic_EOS}
\subfile{Sections/Thermodynamic_details}

### 2023-10-05 SFE_article_Parameter_Estimation/Sections/Thermo_imp.tex
03.10
05.10

	A supercritical fluid (SCF) is a substance at a temperature and pressure above its critical point, where there are no distinct liquid and gas phases but below the pressure required to compress it into a solid. SCFs can move through porous solids like gases, which is faster than liquid transport through such materials. SCFs have a higher ability to dissolve materials like liquids or solids compared to gases. Near the critical point, small changes in pressure or temperature result in significant changes in density, allowing many properties of an SCF to be fine-tuned. By changing the pressure and temperature, the properties can be tuned to be more liquid-like or gas-like.

@@

	Supercritical $CO_2$'s thermodynamic properties, such as density or specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects are presented in Appendix \ref{subsubsec: Equation of state}. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a gas as a function of temperature and density.

	To determine the thermodynamic properties of a real gas, it is necessary to evaluate the departure function of the chosen equation of state for that property. As explained by \citet{Elliott2011}, the departure function is the difference between the actual value of a thermodynamic property of a real gas and its value if the gas were ideal under the same temperature and pressure conditions. The ideal gas serves as a reference state to which the properties of real gases are compared. The departure function measures the extent to which a real gas deviates from ideal gas behaviour. The departure functions allow for the accurate calculation of thermodynamic properties for real gases. %(10.1007/s10494-017-9872-4)

@@

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor, and its value depends on the operating conditions. The fluid properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in the Appendix \ref{subsubsec: Fluid density}.

@@

@@

	Similarly, the NIST developed the correlation, which describe the behaviour of thermal diffusivity of $CO_2$. The work of \citet{Huber2016} captures the singular behaviour of thermal conductivity around the critical point. The correlation is applicable for the temperature range from the triple point to 1100 K and pressures up to 200 MPa. Figure \ref{fig: SFE_Properties_Thermal}. The presented figure show regions around the critical point where the singularity is present.

>>>>

	A supercritical fluid (SCF) is a substance at a temperature and pressure above its critical point, where there are no distinct liquid and gas phases but below the pressure required to compress it into a solid. SCFs can move through porous solids like gases, which is faster than liquid transport through such materials. SCFs have a higher ability to dissolve materials like liquids or solids compared to gases. Near the critical point, small changes in pressure or temperature result in significant changes in density, allowing many properties of an SCF to be fine-tuned. By changing the pressure and temperature, the properties of carbon dioxide can be tuned to be more liquid-like or gas-like.

@@

	Supercritical $CO_2$'s thermodynamic properties, such as density or specific heat capacity, vary significantly with slight changes in temperature and pressure due to real gas effects. The Peng-Robinson equation of state (P-R EOS) is used to calculate the thermodynamic properties by accounting for these real gas effects are presented in Appendix \ref{subsubsec: Equation of state}. The P-R EOS belongs to a specific class of thermodynamic models for modelling the pressure of a solvent as a function of temperature and density and which can be rewritten as a cubic function of the molar volume.

	To determine the thermodynamic properties of a real gas, it is necessary to evaluate the departure function of the chosen equation of state for that property. As explained by \citet{Elliott2011}, the departure function describe the difference between the actual value of a thermodynamic property of a real gas and its value if the gas were ideal under the same temperature and pressure conditions. The ideal gas serves as a reference state to which the properties of real gases are compared. The departure function measures the extent to which a real gas deviates from ideal gas behaviour. The departure functions allow for the accurate calculation of thermodynamic properties for real gases. %(10.1007/s10494-017-9872-4)

@@

	The real gas effects are also visible on the density plot presented in Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor, and its value depends on the operating conditions. The fluid properties near the critical point are unique and combine gas-like and liquid-like properties. The details of calculations are explained in Appendix \ref{subsubsec: Fluid density}.

@@

	Figure \ref{fig: SFE_Properties_h} represent how the specific enthalpy change with the operating conditions. The details of calculations are discussed in Appendix \ref{CH:Enthalpy}.

@@

	Similarly, the NIST developed the correlation, which describe the behaviour of thermal diffusivity of $CO_2$. The work of \citet{Huber2016} captures the singular behaviour of thermal conductivity around the critical point. The correlation is applicable for the temperature range from the triple point to 1100 K and pressures up to 200 MPa. Figure \ref{fig: SFE_Properties_Thermal} shows how the thermal conductivity of carbon dioxide change depending on the operating conditions.
